---
title: "Gapminder Unwinder: Hwk_2"
author: "d8a-m8 a.k.a. delta thru data a.k.a Yevgen 'the d8a nonh8a' Kovalenko"
date: "`r format (Sys.Date())`"
output: github_document
##yamal front matter ^^
---

~~~~~
**again just messing around to make this class more fun and engaging for myself**
~~~~~

## Load Gapminder data

...reflections...
A
```{r setup, include=F}
knitr::opts_chunk$set(echo=TRUE) #above: hides this chunk, inline:sets all chunks to be viewed globally

library('gapminder')
library('tibble')
library('tidyverse')
library('random')
```

```{r}

#Exploring the data frame. You will need internet connection to run some of this code.
gapminder
summary(gapminder[,3:6])
str(gapminder)
```

Let's explore the relationship between life expectancy as a fucntion of GDP per capita.

```{r}
plot(log(lifeExp) ~ log(gdpPercap), data = gapminder)
```

As seen above, there is some correlation between life expectancy and GDP per capita. Perhaps this is trivial becasue as GDP per capita increases, it could be said that quality of life also increases and thus, perhaps life expectancy. Although this clearly does not paint the complete picture as the other factors will affect both that may or may not affect the other.

```{r}
#gapminder...looking at continental graphical distributions of population, life expectancy, and GDP
plot(lifeExp ~ continent, data = gapminder)
plot(log10(gdpPercap) ~ continent, data = gapminder)
```

The above trends may reflect the socio-economic status of a continent as life expectancy can be seen to corelate with GDP. Let's take a look to see if this trend holds up for different places. Let us make an abitrary decision on which countries to investigate.

```{r}
country_names <- unique(gapminder['country'])

cnm <- grep(
  as.list(randomStrings(n=1,len=1,digits=F)), 
  as.matrix(country_names[,1]))

name.list <- country_names[cnm,]

print(name.list)

new_gp <- gapminder %>% 
  filter(country %in% c(as.matrix(name.list)))

new_gp_data <- new_gp %>% 
  select(lifeExp, gdpPercap)
  
print(new_gp)
```

# %>% pipe operator fucn(arg1,arg2...) v. arg1 %>% func(arg2....)
  
```

```{r}

```

 
